Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2023-11-27 article.qmd
Update bibliography for vscode-antimony
Change title and introduction to include poincaré

---
title: "SimBio: a versatile and expandable Python package for modeling systems."
bibliography: bibliography.bib
format:
  html:
    number-sections: true
    anchor-sections: true
  pdf:
    documentclass: scrartcl
    papersize: a4
    geometry:
        - margin=20mm
    pdf-engine: tectonic
    number-sections: true
    colorlinks: true
  docx:
    number-sections: true
---

@@

Another weakness is related to the way models are written.
While most tools support standards such as the System Biology Markup Language (SBML) [@huckaSBMLL3V2],
no one is expected to write an SBML file directly.
Tellurium uses a Domain Specific Language (DSL) called Antimony [@smithAntimonyModularModel2009],
which can be translated to SBML.
A DSL allows code to be reused in different programming environments,
but it will not be recognized by Integrated Development Environments (IDEs)
and therefore cannot provide the development experience we have come to expect, such as
syntax highlighting, code completion, refactoring, and static analysis.
This has been recognized by the community,
and an extension for Visual Studio Code has been created [@vscode-antimony].
However, maintaining it in the future could be burdensome for the smaller systems biology community.
TODO: this is missing for other tools

We developed SimBio,
an open-source Python package for defining and simulating reaction networks,
to address these two issues.
It focuses on providing an ergonomic experience for end-users
by integrating well with IDEs and static analysis tools
through the use of standard modern Python syntax.
It can be extended from Python
since it is coded in pure Python,
which is beneficial for power users.
TODO: high degree of flexibility,  large potential to leverage existing python tools.  For example, ....
Its models can be introspected to create other representations,
such as graphs connecting species and/or reactions,
or tables with parameters or equations.
Finally,
it has a modular architecture with a clear separation of concerns into three packages,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.

>>>>

---
title: "Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems."
bibliography: bibliography.bib
format:
  html:
    number-sections: true
    anchor-sections: true
  pdf:
    documentclass: scrartcl
    papersize: a4
    geometry:
        - margin=20mm
    pdf-engine: tectonic
    number-sections: true
    colorlinks: true
  docx:
    number-sections: true
---

@@

Another weakness is related to the way models are written.
While most tools support standards such as the System Biology Markup Language (SBML) [@huckaSBMLL3V2],
no one is expected to write an SBML file directly.
Tellurium uses a Domain Specific Language (DSL) called Antimony [@smithAntimonyModularModel2009],
which can be translated to SBML.
A DSL allows code to be reused in different programming environments,
but it will not be recognized by Integrated Development Environments (IDEs)
and therefore cannot provide the development experience we have come to expect, such as
syntax highlighting, code completion, refactoring, and static analysis.
This has been recognized by the community,
and an extension for Visual Studio Code has been created [@maTestingDynamicalSystems2021].
However, maintaining it in the future could be burdensome for the smaller systems biology community.
TODO: this is missing for other tools

To address these issues,
we developed poincaré and SimBio,
open-source Python packages for defining and simulating systems.
Poincaré allows to define differential equation systems
and SimBio builds on top of it to define reaction networks.
They are focused on providing an ergonomic experience for end-users
by integrating well with IDEs and static analysis tools
through the use of standard modern Python syntax.
They can be extended from Python
since they are coded in pure Python,
which is beneficial for power users.
TODO: high degree of flexibility,  large potential to leverage existing python tools.  For example, ....
Their models can be introspected to create other representations,
such as graphs connecting species and/or reactions,
or tables with parameters or equations.
Finally,
they have a modular architecture with a clear separation of concerns,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.
